Efficient Handling of String-Number Conversion

### 2019-11-22 evaluation.tex
Update table

We have implemented our string constraint solving procedure, in a tool called {\tool}, on top of the SMT solver Z3~\cite{z3} as a theory solver. In this way we can concentrate on solving constraints connected by conjunctions and let Z3 to handle other boolean connectives. It also enable us to solve not only formulas over string constraints, but also the combination of string constraints with constraints over other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulas.

In {\tool}, we use the following PFA selection strategy: we select PFA's of size $m$ for string variables occur a number-string conversion constraint and PFA's with $p$ loops of size $q$ for other variables. Initially, we set $(m,p,q)=(X,X,X)$, and increase each of them by one if more precise refinement is required. We set an upper bound $7$ for each parameters and report UNKNOWN if a solution cannot be found within the bound. \yfc{need to check with Diep for the parameters}.

@@

We compare {\tool} with other state-of-the-art string solvers, namely CVC4~(relese-1.7)~\cite{cvc4Tool}, Z3~(\texttt{dde8da853})~\cite{z3}, \textsf{Z3-str3}~(\texttt{109dfc5e8})~\cite{zheng2013z3}, and \textsf{Trau+}~\cite{abdulla2019chain}. We do not compare with Sloth \cite{sloth} since it does not support length constraints, which occurs in most of our benchmarks. We also do not compare with ABC~\cite{aydin2018parameterized} (a model counter for string constraints) and Ostrich~\cite{chen2019decision}, because most of our benchmarks contain non-trivial equality constraints, which is not fully supported by those solvers.

We perform two sets of experiments. In the first experiment, we compare with other tools on benchmarks over basic string constraints. We use the following benchmark examples in this experiments:
\begin{itemize}
	\item PyEx~\cite{pyex} is from running the symbolic executor PyEx over some Python packages.
	\item APLAS~\cite{aplas} includes 600 hand-drafted tests consisting of only equality and integer length constraints.
	\item LeetCode is from running PyEx over some sample code collected from the LeetCode~\cite{LeetCode} website.
	\item StringFuzz~\cite{blotsky2018stringfuzz}  is generated from the fuzz testing tool of the same name.
	\item $\text{cvc4}_{\text{pred}}$ and $\text{cvc4}_{\text{term}}$ are obtained from the CVC4 group~\cite{termEQ}. These benchmarks contain a small amount of string-number conversion constraints (< $5\%$).
\end{itemize}

In the second experiment, we compare with tools supporting to string-number conversion on the  benchmarks collected from the symbolic executor \texttt{Py-Conbyte}\footnote{\url{https://github.com/spencerwuwu/py-conbyte}}, which has the supports of string-number conversion. We run it on several examples collected from the LeetCode platform and from Python core libraries. We also have examples that encode execution paths of some JavaScript programs (the Luhn algorithm and some array manipulation).

All experiments were executed on a computer with 4-core CPU, 8 GiB RAM. The timeout was set to 10s for each test.
We use the results from {\tool}, CVC4, and Z3 as the reference answer for the validation of the correctness of the results. Occasionally, two of them report inconsistent (one SAT and one UNSAT) answers to the same test. To decide which solver reports the incorrect answer, we developed a validator, which takes the model $I$ returned from the solver who reported SAT, assigns $I(x)$ to all variables $x$ in the test to obtained a modified test, and re-evaluates the modified test by multiple solvers. If the results from all solvers are consistent, we mark the test SAT or UNSAT according to the results. Otherwise, we manually simplify and inspect the test until we get a conclusive result.

The results of the experiments are summarized in Table~\ref{table:base_benchmark} and Table~\ref{table:str_int_benchmark}, respectively. Columns with heading \texttt{sat}/\texttt{unsat} show numbers of solved formulas. Columns with heading $\times$ indicate the number of instances the solver fails (including timeout, report unknown, report incorrect result, or crash).

\begin{table}[h]
\centering
\caption{Results of {\tool}, CVC4, and Z3/Z3str3 on Basic String Constraint benchmarks.}
\scalebox{0.85}{
\begin{tabular}{|l r | r r r r r|}
\hline
\multicolumn{2}{|c}{}                   & {\tool} & CVC4  &Z3 & Z3-str3 & Trau+  \\ \hline
\multirow{3}{*}{PyEx}		& SAT      & 19586*   & 19763 & 18490 &  3514 & 19149 \\ 
							& UNSAT    &  3858*   &  3834 &  3855 &  4521 & 3828 \\
							& $\times$ &  1977*   &  1824 &  3076 & 17386 & 2444 \\ \hline
\multirow{3}{*}{APLAS}		& SAT      &   128*   &   205 &    19 &    52 & 132 \\
							& UNSAT    &   287*   &   221 &   100 &   118 & 82 \\
							& $\times$ &   185*   &   174 &   481 &   430 & 386 \\ \hline
\multirow{3}{*}{LeetCode}	& SAT      &   856*   &   859 &   881 &   651 & 778 \\
							& UNSAT    &  1784*   &  1785 &  1785 &  1787 & 1827 \\
							& $\times$ &    26*   &    22 &     0 &   228 & 61 \\ \hline
\multirow{3}{*}{StringFuzz}	& SAT      &   502*   &   671 &   316 &   605 & 688 \\
							& UNSAT    &   294*   &   240 &   206 &   217 & 339 \\
							& $\times$ &   269*   &   154 &   543 &   243 & 38 \\\hline
\multirow{3}{*}{cvc4\textsubscript{pred}} & SAT & 16* & 11 &   12 &    10 & 247 \\
							& UNSAT    &   814*   &   818 &   807 &   778 & 500 \\
							& $\times$ &     5*   &     6 &    16 &    47 & 88 \\ \hline
\multirow{3}{*}{cvc4\textsubscript{term}} & SAT & 13* & 8 &     6 &    16 & 311 \\
							& UNSAT    &  1030*   &   936 &  1020 &   962 & 596 \\
							& $\times$ &     2*   &   101 &    19 &    67 & 138 \\ \hline
\multirow{3}{*}{Total} 		& SAT      & 21101*  &  21517 & 19724 &  4848 & 21305  \\
							& UNSAT    &  8067*  &   7834 &  7773 &  8383 & 7172  \\
							& $\times$ &  2464*  &   2281 &   4135 & 18401 & 3155  \\\hline	
\end{tabular}}
\label{table:base_benchmark}
\end{table}

\begin{table}[h]
\centering
\caption{Results of {\tool}, CVC4, and Z3/Z3str3 on Basic String Constraints + String-Number Conversion benchmark.}
\scalebox{0.8}{
\begin{tabular}{|l r | r r r r|}
\hline
\multicolumn{2}{|c}{}                   			& {\tool} & CVC4   &    Z3  & Z3-str3 \\ \hline
\multirow{3}{*}{String-Number-Leetcode}	 & SAT      &   2356*  &  1543 &  1025  &   333 \\ 
										 & UNSAT    &  16369*  & 15676 & 17234  & 15543 \\
										 & $\times$ &    202*  &  1708 &   668  &  3051 \\ \hline 
\multirow{3}{*}{String-Number-PythonLib} & SAT      &    912*  &   408 &  1802  &   389 \\ 
										 & UNSAT    &    720*  &   660 &   744  &   729 \\
										 & $\times$ &   1014*  &  1578 &   100  &  1528 \\ \hline
\multirow{3}{*}{String-Number-JavaScript}& SAT      &     20   &     6 &    16  &     4 \\ 
										 & UNSAT    &      0   &     0 &     0  &     0 \\
										 & $\times$ &      0   &    14 &     4  &    16 \\ \hline \hline
\multirow{3}{*}{Total}					 & SAT      &   3288*  &  1957 &  2843  &   726 \\ 
										 & UNSAT    &  17089*  & 16336 & 17978  & 16272 \\
										 & $\times$ &   1216*  &  3300 &   772  &  4595 \\ \hline
\end{tabular}}
\label{table:str_int_benchmark}
\end{table}

From Table~\ref{table:base_benchmark}, we can see that the performance  of {\tool} is as good as the most competitive tools such as CVC4 and Z3 on basic string constraints. In all of the benchmarks, {\tool} ranked either the 1st or the 2nd on the number of solved (SAT+UNSAT) cases. If we look into scrutiny, for the APLAS and StringFuzz benchmarks, there is a significant distance between {\tool} and the top-ranked tool on the SAT cases. But we believe this is not very important because both these two cases are artificial, either hand-crafted or randomly produced from a fuzzer. The more critical ones are those come from program analysis, which {\tool} has comparable performance to the best tool.

From Table~\ref{table:str_int_benchmark}, we can see that {\tool} significantly outperforms all other tools. The total number of failed cases is only a half to Z3, which is ranked the 2nd. In fact, most of the failed tests comes from the analysis of one Python core library function that converts an IPv6 address to a number. We generated 2028 tests from this function, and among them around 1000 cases that are too difficult for all solvers.  If we exclude results on this function, then {\tool} has only $185$ failed case in total. This is significantly better than the 2nd place tool Z3, which fails $893$ cases in total.

>>>>

We have implemented our string constraint solving procedure in a tool called {\tool}. {\tool} is implemented on top of the SMT solver Z3~\cite{z3} as a theory solver. In this way we can concentrate on solving constraints connected by conjunctions and let Z3 to handle other boolean connectives. It also enables us to solve not only formulae over string constraints, but also the combination of string constraints with constraints over other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulae.

In {\tool}, we use the following PFA selection strategy: we select PFAs of size $m$ for string variables involved in a number-string conversion constraint and PFAs with $p$ loops of size $q$ for other variables. Initially, we set $(m,p,q)=(X,X,X)$, and increase each of them by one if more precise refinement is required. We set an upper bound $7$ for each parameters and report UNKNOWN if a solution cannot be found within the bound. \yfc{need to check with Diep for the parameters}.

@@

We compare {\tool} with other state-of-the-art string solvers, namely CVC4~(relese-1.7)~\cite{cvc4Tool}, Z3~(\texttt{dde8da853})~\cite{z3}, \textsf{Z3-str3}~(\texttt{1a9dfc5e8})~\cite{zheng2013z3}, and \textsf{Trau+}~\cite{abdulla2019chain}. We do not compare with Sloth \cite{sloth} since it does not support length constraints, which occurs in most of our benchmarks. We also do not compare with ABC~\cite{aydin2018parameterized} (a model counter for string constraints) and Ostrich~\cite{chen2019decision}, because most of our benchmarks contain non-trivial equality constraints, which is not fully supported by those solvers.

We perform two sets of experiments. In the first set of experiment, we compare with other tools on benchmarks over basic string constraints. We use the following benchmark examples in this experiments:
\begin{itemize}
	\item PyEx~\cite{pyex} is from running the symbolic executor PyEx over some Python packages.
	\item APLAS~\cite{aplas} includes 600 hand-drafted tests consisting of only equality and integer length constraints.
	\item LeetCode is from running PyEx over some sample code collected from the LeetCode~\cite{LeetCode} website.
	\item StringFuzz~\cite{blotsky2018stringfuzz}  is generated from the fuzz testing tool of the same name.
	\item $\text{cvc4}_{\text{pred}}$ and $\text{cvc4}_{\text{term}}$ are obtained from the CVC4 group~\cite{termEQ}. These benchmarks contain a small amount of string-number conversion constraints (< $5\%$).
\end{itemize}

In the second experiment, we compare with tools supporting string-number conversion on the benchmarks collected from the symbolic executor \texttt{Py-Conbyte}\footnote{\url{https://github.com/spencerwuwu/py-conbyte}}, which has the supports of string-number conversion. We run it on several examples collected from the LeetCode platform and from Python core libraries. We also have examples that encode execution paths of some JavaScript programs (the Luhn algorithm and some array manipulation).

All experiments were executed on a machine with 4-core CPU, 8 GiB RAM. The timeout was set to 10s for each test.
We use the results from {\tool}, CVC4, and Z3 as the reference answer for the validation of the correctness of the results. Occasionally, two of them report inconsistent (one SAT and one UNSAT) answers to the same test. To decide which solver reports the incorrect answer, we developed a validator, which takes the model $I$ returned from the solver who reported SAT, assigns $I(x)$ to all variables $x$ in the test to obtained a modified test, and re-evaluates the modified test by multiple solvers. If the results from all solvers are consistent, we mark the test SAT or UNSAT according to the results. Otherwise, we manually simplify and inspect the test until we get a conclusive result.

The results of the experiments are summarized in Table~\ref{table:base_benchmark} and Table~\ref{table:str_int_benchmark}, respectively. Columns with heading \texttt{SAT}/\texttt{UNSAT} show numbers of solved formulae. Rows with heading $\times$ indicate the number of instances the solver fails (including timeout, report unknown, report incorrect result, or crash).

\begin{table}[h]
\centering
\caption{Results of {\tool}, CVC4, and Z3/Z3str3 on Basic String Constraint benchmarks.}
\scalebox{0.85}{
\begin{tabular}{|l r | r r r r r|}
\hline
\multicolumn{2}{|c}{}                  & {\tool} & CVC4  &Z3 & Z3-str3 & Trau+  \\ \hline
\multirow{3}{*}{PyEx}		& SAT      &  21205  & 19763 & 18490 &  3514 & 19149 \\ 
							& UNSAT    &   3963  &  3834 &  3855 &  4521 & 3828 \\
							& $\times$ &    253  &  1824 &  3076 & 17386 & 2444 \\ \hline
\multirow{3}{*}{APLAS}		& SAT      &    135  &   205 &    19 &    52 & 132 \\
							& UNSAT    &    343  &   221 &   100 &   118 & 82 \\
							& $\times$ &    122  &   174 &   481 &   430 & 386 \\ \hline
\multirow{3}{*}{LeetCode}	& SAT      &    842  &   859 &   881 &   651 & 778 \\
							& UNSAT    &   1824  &  1785 &  1785 &  1787 & 1827 \\
							& $\times$ &      0  &    22 &     0 &   228 & 61 \\ \hline
\multirow{3}{*}{StringFuzz}	& SAT      &    528  &   671 &   316 &   605 & 688 \\
							& UNSAT    &    320  &   240 &   206 &   217 & 339 \\
							& $\times$ &    217  &   154 &   543 &   243 & 38 \\\hline
\multirow{3}{*}{cvc4\textsubscript{pred}} & SAT & 13 & 11 &   12 &    10 & 247 \\
							& UNSAT    &    820  &   818 &   807 &   778 & 500 \\
							& $\times$ &      2  &     6 &    16 &    47 & 88 \\ \hline
\multirow{3}{*}{cvc4\textsubscript{term}} & SAT & 13 & 8 &     6 &    16 & 311 \\
							& UNSAT    &   1031  &   936 &  1020 &   962 & 596 \\
							& $\times$ &      1  &   101 &    19 &    67 & 138 \\ \hline
\multirow{3}{*}{Total} 		& SAT      &  22736  &  21517 & 19724 &  4848 & 21305  \\
							& UNSAT    &   8301  &   7834 &  7773 &  8383 & 7172  \\
							& $\times$ &    595  &   2281 &   4135 & 18401 & 3155  \\\hline	
\end{tabular}}
\label{table:base_benchmark}
\end{table}

\begin{table}[h]
\centering
\caption{Results of {\tool}, CVC4, and Z3/Z3str3 on Basic String Constraints + String-Number Conversion benchmark.}
\scalebox{0.8}{
\begin{tabular}{|l r | r r r r|}
\hline
\multicolumn{2}{|c}{}                   			& {\tool} & CVC4   &    Z3  & Z3-str3 \\ \hline
\multirow{3}{*}{String-Number-Leetcode}	 & SAT      &   2413  &  1543 &  1025  &   333 \\ 
										 & UNSAT    &  16511  & 15676 & 17234  & 15543 \\
										 & $\times$ &      3  &  1708 &   668  &  3051 \\ \hline 
\multirow{3}{*}{String-Number-PythonLib} & SAT      &    424  &   408 &  1802  &   389 \\ 
										 & UNSAT    &   2056  &   660 &   744  &   729 \\
										 & $\times$ &    166  &  1578 &   100  &  1528 \\ \hline
\multirow{3}{*}{String-Number-JavaScript}& SAT      &     20  &     6 &    16  &     4 \\ 
										 & UNSAT    &      0  &     0 &     0  &     0 \\
										 & $\times$ &      0  &    14 &     4  &    16 \\ \hline \hline
\multirow{3}{*}{Total}					 & SAT      &   2857  &  1957 &  2843  &   726 \\ 
										 & UNSAT    &  18567  & 16336 & 17978  & 16272 \\
										 & $\times$ &    169  &  3300 &   772  &  4595 \\ \hline
\end{tabular}}
\label{table:str_int_benchmark}
\end{table}

From Table~\ref{table:base_benchmark}, we can see that the performance of {\tool} is as good as the most competitive tools such as CVC4 and Z3 on basic string constraints. In all of the benchmarks, {\tool} ranked either the 1st or the 2nd on the number of solved (SAT+UNSAT) cases. If we look into scrutiny, for the APLAS and StringFuzz benchmarks, there is a significant distance between {\tool} and the top-ranked tool on the SAT cases. But we believe this is not very important because both these two benchmarks are artificial, either hand-crafted or randomly produced from a fuzzer. The more critical ones are those come from program analysis, which {\tool} has comparable performance to the best tool.

From Table~\ref{table:str_int_benchmark}, we can see that {\tool} significantly outperforms all other tools. The total number of failed cases is only one-fourth to Z3, which is ranked the 2nd. In fact, most of the failed tests comes from the analysis of one Python core library function that converts an IPv6 address to a number. We generated 2028 tests from this function, and among them around 150 cases that are too difficult for all solvers.  If we exclude results on this function, then {\tool} has only $185$ failed case in total. This is significantly better than the 2nd place tool Z3, which fails $893$ cases in total. \yfc{This paragraph needs to be revised when new data of Z3-Trau is available. Note that z3seq is much more better now. Also is z3-trau.}

\hide{ % data for helping to write comparison
benchmark_path: lib_int-ipaddress__ip_int_from_string
benchmark_size: 2028
(tool-check_date, #sat, #unsat, #timeout, #unknown, #exception/error, #misc)
(cvc4-20191120, 96, 367, 1565, 0, 0, 0)
(z3-trau-20191122, 135, 1739, 154, 0, 0, 0)
(z3seq-20191118, 1509, 431, 88, 0, 0, 0)
(z3str3-20191118, 191, 440, 1330, 5, 0, 62)

benchmark_path: leetcode_int-
benchmark_size: 18927
(tool-check_date, #sat, #unsat, #timeout, #unknown, #exception/error, #misc)
(z3seq-20191118, 1025, 17234, 668, 0, 0, 0)
(cvc4-20191120, 1543, 15676, 1708, 0, 0, 0)
(cvc4-20191118, 1766, 15612, 1549, 0, 0, 0)
(z3-trau-20191122, 2413, 16511, 3, 0, 0, 0)
(z3str3-20191118, 333, 15543, 2031, 952, 0, 68)

benchmark_path: lib_int-
benchmark_size: 2646
(tool-check_date, #sat, #unsat, #timeout, #unknown, #exception/error, #misc)
(cvc4-20191118, 586, 666, 1394, 0, 0, 0)
(z3seq-20191118, 1802, 744, 100, 0, 0, 0)
(cvc4-20191120, 408, 660, 1578, 0, 0, 0)
(z3str3-20191118, 389, 729, 1392, 70, 0, 66)
(z3-trau-20191122, 424, 2056, 166, 0, 0, 0)
}
